Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2019-04-29 Paper/sync.tex
Updates from Overleaf

We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models. We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits. At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.  In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}. Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the \xxx{Constant Phase Lag tidal} model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d. Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology \xxx{ages can err by up to $300\%$} unless one accounts for binarity. We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.

@@

We examine stellar angular momentum evolution in low-mass binaries by simulating four sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism, using both magnetic braking models.  We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e. more massive star in binaries, as it is observationally easier to measure a P$_{rot}$ on the more massive, and hence brighter, star \citep[e.g.][]{Meibom2006,Lurie2017}. For each simulation, we sample the primary's mass uniformly over $[0.1, 1]$ M$_{\odot}$. Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old Orion Nebula Cluster \citep{Stassun1999,Herbst2001,Herbst2002,Rodriguez-Ledesma2009}.  We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}. Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}. Although the CTL model is applicable for $e \gsim 0.3$, \xxx{the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}}, so we restrict $e \leq 0.3$ to allow us to compare both models.
We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.

Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain due to complex viscous evolution within the stars \citep{Ogilvie2007}, and can differ for stars of the same spectral class \citep{Barker2009}. These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}. Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, \xxx{respectively} \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.  Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^8]$ and $\tau$ log-uniformly over $[10^{-2},10]$ s.  There is no general expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g. Eqn. (2) from \citet{Heller2011}. \xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.}

We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them. All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.

@@

In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation. In this case when P$_{rot} >$ P$_{eq}$ for long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}. Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g. Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as the stellar rotation slows down, so at some P$_{rot}$, tidal torques that try to spin-up subsynchronous rotators back towards P$_{eq}$ will balance the spin-down torque of magnetic braking, producing a long-lasting state of subsynchronous rotation.

We compute the P$_{rot}$ at which this balance occurs by by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.  In this calculation, we consider tidal torques under the CTL formalism and magnetic braking under the \citet{Matt2015} model. For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly, allowing us to set $R = 1 R_{\odot}$ and assume constant moments of inertia. We find below in simulations that these assumptions are well-justified. Given these assumptions, we can specify P$_{orb}$, $k_2$, and $\tau$ to compute the equilibrium P$_{rot}$ at which tidal and magnetic braking torques balance.  We derive this balance, given by Eqn.~\ref{eqn:appendixBalance}, in $\S$~\ref{sec:appendix:balance} and visualize it in Fig.~\ref{fig:analyticBalance}. We normalize P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$ (see Eqn.~(\ref{eqn:ctl:eqPer})).

\begin{figure}[h]
	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
   \caption{The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}. The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.}%
    \label{fig:analyticBalance}%
\end{figure}

In general as tides weaken, i.e. increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where magnetic braking torques are weaker, or equivalently at larger P$_{rot}/$P$_{eq}$. Conversely, in the strong tides regime, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation. Given this calculation, we expect that for many orbital periods and tidal time lags, stellar binaries will reach a state of subsynchronous rotation. For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries on low eccentricity orbits reach the subsynchronous state with P$_{rot}/$P$_{eq} \approx 1.1-1.4$, with this value increasing with P$_{orb}$. This simple analytic calculation, however, does not account for stellar evolution or secular orbital evolution due to tides, e.g. tidal friction that will shrink the orbit, gradually strengthening tidal torques, so we cannot precisely determine when the binaries reach the subsynchronous state nor how long it will last. For that, we turn to simulations.

Below, we verify the results of our analytic calculations by examining how stellar binaries evolve towards subsynchronous rotation, under what conditions, and when it occurs in the weak tides regime. We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a combination of both tidal models and both magnetic braking models. In Fig.~\ref{fig:eqPer}, we plot the simulated evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models.

@@

Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}. The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems. The CTL model, however, predicts that for P$_{orb} \gsim 10$ d, magnetic braking overpowers weak tidal torques, spinning down the stars past the tidally-locked state. Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$ with the maximum value set by the torque balance between tidal torques and magnetic braking, in good agreement with the analytic calculations in $\S$~\ref{sec:analytic}. As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g. Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution, driving the binaries towards subsynchronous rotation.

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model. Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d. The black dotted line indicates the tidally-locked state where P$_{rot} = $ P$_{eq}$.  Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.  We truncate each curve when the binary tidally locks. For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.  For all P$_{orb}$, the CPL models tidally-lock into synchronous rotation whereas the CTL model predicts subsynchronous (P$_{rot} > $ P$_{eq}$) rotation that persists for Gyrs.}%
    \label{fig:eqPer}%
\end{figure*}

@@

Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.  First, as P$_{rot}$ continue to increase due to magnetic braking, magnetic braking itself weakens as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}.  Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega}), for example). Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$ and strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.  These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long term.

We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to stellar evolution, magnetic braking, and tidal torques as a function of time in our simulations. Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the main sequence. After $R$ contraction, $\dot{\mathrm{P}}_{rot} > 0$ as tides and magnetic braking combine to spin down stars towards the tidally-locked state. For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ driving the stars towards subsynchronous rotation. $\ddot{\mathrm{P}}_{rot} < 0$, however, as the three processes described above gradually strengthen tidal torques relative to magnetic braking.  In the long-run, tidal torques overpower magnetic braking, with a slightly negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$ (see the P$_{orb} = 20, 40$ d cases in Fig.~\ref{fig:eqPer}). This slowly shifting torque balance produces a population of subsynchronous rotators with P$_{rot} > $ P$_{eq}$ that can persist for Gyrs.

@@

Spin-orbit synchronization is the typical outcome for binaries with P$_{orb} < 10$ d according to the CPL model for most values of $Q$. The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to P$_{orb} \approx 100 $ d for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.  According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.  Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e. stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker tidal torques allowing more subsynchronous rotation.  The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.

@@

\xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}. For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems, especially for P$_{orb} > 20$ d. For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for P$_{orb} < 70$ d as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e. Eqn.~(\ref{eqn:cpl:eqPer}).

Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking that spin-down the stars past the tidally-locked state.  This effect can be seen for P$_{orb} < 50$ d where the median $Q$ gradually increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.  This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.  In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be.

@@

According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d. \xxx{for P$_{orb} < 50$ d, tidal-locking can occur when $\tau \gsim 0.25$ s  whereas subsynchronous rotation occurs for stars with $\tau < 0.1$ s}.  For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tides given the wide orbital separations}.  The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution. We do not \xxx{often} observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).

@@

Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster (median CPL, CTL P$_{rot} = 27.6$ d and $7.4$ d) than unlocked (median CPL, CTL P$_{rot} = 41.3$ d and $40.8$ d) binaries. More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.  This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model. We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}

>>>>

We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models. We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits. At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.  In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}. Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the \xxx{Constant Phase Lag tidal} model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d. Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology \xxx{can systematically underpredict ages by up to $300\%$} unless one accounts for binarity. We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.

@@

We examine stellar angular momentum evolution in low-mass binaries by simulating \xxx{two} sets of 10,000 stellar binaries, one modeled using the CPL model and the other using the CTL formalism.  We simulate both stars' spin evolution but mainly consider the P$_{rot}$ evolution for the primary, i.e. more massive star in binaries, as it is observationally easier to measure a P$_{rot}$ on the more massive, and hence brighter, star \citep[e.g.][]{Meibom2006,Lurie2017}. For each simulation, we sample the primary's mass uniformly over $[0.1, 1]$ M$_{\odot}$. Following \citet{Matt2015}, we uniformly sample the log$_{10}$ of P$_{rot}$ over [$0.8,15$] days, a distribution that approximates the P$_{rot}$ distribution of young stars in the ${\sim}2$ Myr old Orion Nebula Cluster \citep{Stassun1999,Herbst2001,Herbst2002,Rodriguez-Ledesma2009}.  We compute the secondary star's mass by uniformly sampling the mass ratio over [$0.1, 1$] following observations of mass ratios in low-mass binaries \citep{Raghavan2010,Moe2018}. Given the inherent uncertainty in and complexity of the formation of short-period binaries \citep[e.g.][]{Bonnell1994,Bate2000,Bate2002,Moe2018} and the potential for dynamical processing via tides or stellar close encounters \citep[e.g.][]{Mardling2001,Hurley2002,Ivanova2005,Meibom2005}, we take an agnostic approach to the initial orbital configuration by uniformly randomly sampling the initial eccentricity ($e$) over [$0.0,0.3$], consistent with eccentricities of field binaries that likely have not been tidally-processed \citep{Raghavan2010}. Although the CTL model is applicable for $e \gsim 0.3$, \xxx{the CPL model is not and can predict qualitatively incorrect evolution in that regime \citep[see Section 4.1 in][]{Leconte2010}}, so we restrict $e \leq 0.3$ to allow us to compare both models. We uniformly sample the initial P$_{orb}$ over [$3,100$] d and do not consider P$_{orb} < 3$ d as these binaries are likely to have a tertiary companion \citep{Tokovinin2006} which can significantly impact the inner binary's dynamical evolution \citep[e.g.][]{Munoz2015,Martin2015b,Hamers2016,Moe2018}.

Values for stellar tidal $Q$s and $\tau$s for low-mass stars are highly uncertain due to complex viscous evolution within the stars \citep{Ogilvie2007}, and can differ for stars of the same spectral class \citep{Barker2009}. These parameters can also vary as a function of stellar mass or age \citep{Bolmont2016,vanEylen2016}, likely due to low-mass stars' evolving convective regions where the tidal dissipation predominantly occurs \citep{Zahn2008}. Typical values of $Q$ and $\tau$ for Sun-like stars are estimated to be of order $Q \approx 10^6$ and $\tau \approx 0.1$ s, \xxx{respectively} \citep[e.g.][]{Meibom2005,Ogilvie2007,Jackson2008}, however a range of values exist in the literature.  Therefore, we consider a wide range of tidal parameters by sampling stellar tidal $Q$s log-uniformly over $[10^4,10^8]$ and $\tau$ log-uniformly over $[10^{-2},10]$ s.  There is no general expression to compute $Q$ as a function of $\tau$, or vice versa, except in some special cases where approximations exist, e.g. Eqn. (2) from \citet{Heller2011}. \xxx{All stars have an initial age of 5 Myr unless stated otherwise as by this time, the gaseous protoplanterary circumbinary disk that can drive significant dynamical evolution in the binary \citep[e.g.][]{Fleming2017} would likely have dissipated \citep{Haisch2001}.} We also perform a smaller subset of simulations to illustrate the behaviour of our coupled model and describe their initial conditions as we introduce them. All code used to run simulations and generate figures is available online\footnote{\href{https://github.com/dflemin3/sync}{https://github.com/dflemin3/sync}.}.

@@

In the weak tides regime, spin-down due to magnetic braking will drive the stellar P$_{rot}$ past P$_{eq}$, resulting in subsynchronous rotation\xxx{, P$_{rot} >$ P$_{eq}$. For long P$_{orb}$, the stars will be slowly-rotating and in the unsaturated magnetic braking regime \citep{Matt2015}. Since magnetic braking scales as P$_{rot}^{-3}$ for unsaturated rotators, e.g. Eqn.~(\ref{eqn:mattUnSat}), magnetic braking torques weaken as the stellar rotation slows down, so at some P$_{rot}$, tidal torques will balance magnetic braking, producing a long-lasting state of subsynchronous rotation. We compute the P$_{rot}$ at which this balance occurs as a function of P$_{orb}$, $k_2$, and $\tau$ in $\S$~\ref{sec:appendix:balance} by setting the sum of Eqn.~(\ref{eqn:ctl:omega}) and Eqn.~(\ref{eqn:mattUnSat}) equal to 0, considering tidal torques under the CTL formalism. For simplicity, we assume both stars are solar-mass with 0 obliquity, a circular binary orbit, and that the torque balance occurs while the stars are on the main sequence where stellar properties change slowly. Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful pedagogical case to examine here and in simulations below.  We display the results of this calculation in Fig.~\ref{fig:analyticBalance}, normalizing P$_{rot}$ by P$_{eq}$, which for binary stars with 0 obliquity on circular orbits is simply P$_{orb}$.}

\begin{figure}[h]
	\includegraphics[width=0.5\textwidth]{../Plots/analyticTorque.pdf}
   \caption{\xxx{Updated Figure} The stellar P$_{rot}$, normalized by P$_{eq}$, at which the torques due to magnetic braking and tides balance for a 1 M$_{\odot} - 1$ M$_{\odot}$ binary on a circular orbit according to Eqn.~\ref{eqn:appendixBalance}. The white dashed line indicates our fiducial values for $k_2$ and $\tau$, 0.5 and 0.1 s, respectively, that we adopt in the simulations in $\S$~\ref{sec:mattBalance}.}%
    \label{fig:analyticBalance}%
\end{figure}

\xxx{In general as tides weaken, i.e. increasing P$_{orb}$ and/or decreasing $k_2 \tau$, tidal and magnetic braking torques balance at longer P$_{rot}$ where tides are weaker. Conversely, for strong tides, $k_2 \tau \gsim -1$, tidal torques overpower magnetic braking for most orbital periods, tidally-locking binaries into synchronous rotation. For our fiducial values of $k_2 = 0.5$ and $\tau = 0.1$ s (white dashed line in Fig.~\ref{fig:analyticBalance}), solar-twin binaries will rotate subsynchronously for P$_{orb} \gsim 20$ d, with more severe subsynchronism for longer P$_{orb}$. This simple calculation, however, does not account for stellar evolution or secular tidal orbital evolution, e.g. tidal friction that will shrink the orbit, gradually strengthening tidal torques.}

% cut
% , noting that $d\omega/dt = (1/\mathrm{I}) dJ/dt$ for fixed moment of inertia, I.  In this calculation, we

@@

\xxx{We simulate the full coupled stellar-tidal evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits to examine how stellar binaries evolve towards subsynchronous rotation. In Fig.~\ref{fig:eqPer}, we plot P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) models. Both tidal models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}. The CPL model predicts that all binaries tidally-lock, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems. As anticipated by our analytic calculations, the CTL model predicts subsynchronous rotation for P$_{orb} \geq 20$ d as magnetic braking overpowers tidal torques. For P$_{orb} = 20$ d, Magnetic braking pushes P$_{rot}/$P$_{eq} \approx 1.05$, with the maximum value set by the torque balance. As shown in $\S$~\ref{sec:analytic}, the peak P$_{rot}/$P$_{eq}$ grows for longer P$_{orb}$ since tides weaken with increasing binary separation, e.g. Eqn.~(\ref{eqn:ctl:z}), allowing magnetic braking to dominate the spin evolution.}

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelMatt.pdf}
   \caption{\xxx{Updated Figure} Evolution of stellar P$_{rot}$, normalized by P$_{eq}$ (see Eqn.~(\ref{eqn:cpl:eqPer}) and Eqn.~(\ref{eqn:ctl:eqPer}), for initial circular binary orbits according to the CPL (solid) and CTL (dashed) models with $Q = 10^6$ and $\tau = 0.1$ s, respectively, using the \citet{Matt2015} magnetic braking model. Left: P$_{rot}/$P$_{eq}$ for stars with P$_{orb}$ ranging from 5 d to 60 d. The black dotted line indicates the tidally-locked state.  Right: Net P$_{rot}$ derivative due to stellar evolution, tidal torques, and magnetic braking.  We truncate each curve when the binary tidally locks. For P$_{orb} \leq 10$ d, both the CPL and CTL models predict that the binaries lock into synchronous rotation.  For all P$_{orb}$, the CPL models tidally-lock whereas the CTL model predicts subsynchronous rotation that persists for Gyrs.}%
    \label{fig:eqPer}%
\end{figure*}

@@

Subsynchronous rotation does not persist indefinitely, however, as P$_{rot}$ eventually decreases back towards the tidally-locked state in the long-term due to a combination of three simultaneous physical effects.  First, \xxx{magnetic braking weakens at long P$_{rot}$ as its torque scales as P$_{rot}^{-3}$ for unsaturated rotators \citep{Matt2015}. Second, as P$_{rot}$ increases further from the tidally-locked state, tidal torques strengthen as they try to force P$_{rot}$ back towards P$_{eq}$ (see Eqn.~(\ref{eqn:ctl:omega})). Third, when P$_{rot} > $ P$_{eq}$, tides transfer angular momentum from the orbit into stellar rotations, decreasing P$_{orb}$, strengthening tidal torques that strongly depend on the binary separation as $a^{-6.5}$.  These effects combine to shift the balance of power from magnetic braking-controlled stellar spin down to tidal torques spinning-up stars, shepherding them towards P$_{eq}$ in the long-term.}

\xxx{We can see this process unfold in the right panel of Fig.~\ref{fig:eqPer} where we plot the total P$_{rot}$ time derivative due to tidal torques, stellar evolution, and magnetic braking. Early on, $\dot{\mathrm{P}}_{rot} < 0$ as stars contract along the pre-main sequence until about 60 Myr when the stars reach the zero age main sequence. Tides and magnetic braking then combine to spin down stars towards the tidally-locked state. For the CTL models with P$_{orb} > 10$ d, $\dot{\mathrm{P}}_{rot} > 0$ as magnetic braking dominates, driving the stars into subsynchronous rotation. In the long-term, $\ddot{\mathrm{P}}_{rot} < 0$, however, as the three processes described above gradually strengthen tidal torques relative to magnetic braking.  Tidal torques eventually overpower magnetic braking, seen as a slight negative P$_{rot}$ derivative, slowly driving P$_{rot}$ back towards P$_{eq}$, producing a population of subsynchronous rotators that can persist for Gyrs. We explore this point further in $\S$~\ref{sec:shortPorbSubSync}.}

@@

Spin-orbit synchronization is the typical outcome for binaries with P$_{orb} < 10$ d according to the CPL model for most values of $Q$. The strong tidal torques predicted by the CPL model can even tidally-lock binaries out to \xxx{P$_{orb} \gsim 80 $ d} for $Q < 10^5$, well beyond the expected limit of 20 d \citep{Meibom2006}.  According to the CTL model, binaries with P$_{orb} < 10$ d typically tidally-lock for $\tau \gsim 0.1$ s, and seldomly tidally-lock for P$_{orb} > 20$ d, except for systems with strong tides, $\tau \gsim 3$ s.  Both models predict a substantial population of subsynchronous rotators (red regions in Fig.~\ref{fig:qTauLock}, P$_{rot} >$ P$_{orb}$), consistent with magnetic braking dominating weak tidal torques, i.e. stars with large $Q$ or small $\tau$, that form via the mechanism described in $\S$~\ref{sec:eq}, with weaker tidal torques allowing more subsynchronous rotation.  The population of supersynchronous rotators (blue regions in Fig.~\ref{fig:qTauLock}, P$_{rot} <$ P$_{orb}$) with P$_{orb} > 80$ d does not in general correspond to binaries tidally-locking into supersynchronous rotation, but rather, typically arises from the combination of weak tidal torques and magnetic braking not spinning down stars enough for P$_{rot}$ to be close to the tidally-locked state.

@@

\xxx{We isolate the impact of $Q$ and $\tau$ by binning our simulation results after 7 Gyr of evolution by P$_{orb}$ and P$_{orb}/$P$_{rot}$}. For all P$_{orb}$ in Fig.~\ref{fig:qmap}, synchronous and supersynchronous rotators have systematically low $Q$s, typically $Q < 10^6$, as strong tidal torques are required to tidally-lock the systems. For P$_{orb} < 10$ d, there are no rotators with $1.0 <$ P$_{orb}/$P$_{rot} < 1.5$ nor do any stars have P$_{orb}/$P$_{rot} > 1.5$ for \xxx{P$_{orb} < 60$ d} as in the CPL model, binaries with eccentric orbits can only tidally-lock into a 1:1 or 3:2 spin-orbit commensurability, i.e. Eqn.~(\ref{eqn:cpl:eqPer}).

Subsynchronous rotators have systematically larger $Q$s, typically $Q > 10^6$, and hence experience weak tidal torques that are dominated by magnetic braking.  This effect can be seen for P$_{orb} < 50$ d where the median $Q$ \xxx{tends to} increases with decreasing P$_{orb}/$P$_{rot}$, except near the tidally-locked state.  This trend reverses at longer P$_{orb} > 60$ d where supersynchronous rotation arises from the inability of tidal torques and magnetic braking to spin-down stars enough to approach the tidally-locked state by the end of the simulation.  In this case, the more supersynchronous the rotation, the weaker the tidal torques must be, and hence the larger the $Q$ must be.

@@

\xxx{According to the CTL model, many binaries tidally-lock for P$_{orb} \lsim 20$ d when $\tau \gsim 0.1$ s, with some tidally-locking up to P$_{orb} \approx 50$ d when $\tau \gsim 1$ s. Subsynchronous rotation typically occurs for stars with $\tau < 0.1$ s}.  For P$_{orb} > 50$ d, magnetic braking dominates the evolution as the diagonal sequence with a median $\tau \approx 0.1$ s\xxx{, typically insufficient for tide to strongly influence the evolution given the wide orbital separations}.  The shape of this diagonal region arises from the combination of magnetic braking and our flat initial P$_{orb}$ distribution. We do not \xxx{often} observe P$_{orb}/$P$_{rot} \gsim 1.5$ as we only consider eccentricities up to $e = 0.3$, limiting how rapid supersynchronous systems can rotate according to Eqn.~(\ref{eqn:ctl:eqPer}).

@@

Both models predict a substantial population of tidally-locked fast rotators with P$_{rot} \lsim 20$ d, with tidally-locked stars systematically rotating faster \xxx{(median CPL, CTL P$_{rot} = 22.6$ d and $8.8$ d)} than unlocked \xxx{(median CPL, CTL \xxx{both} P$_{rot} = 32.4$ d)} binaries. More massive stars are more likely to tidally-lock compared to less massive stars as tidal torques scale \xxx{with the stellar masses and as $R^5$, with $R$ increasing with stellar mass.  This feature is seen in the enhanced density of locked systems at larger masses for both tidal models, but in particular for the CPL model. We highlight this enhanced density of locked binaries in systems with more massive primaries in the marginalized mass distributions in the top panels of Fig.~\ref{fig:lockedCPL} and Fig.~\ref{fig:lockedCTL}.}
